The persuasion of symbolic elements in government public communications, and their moderation by personal relevance

### 2018-07-04 paper3_relevance and symbols_01.Rmd
Add files via upload

# Methodology

@@

My outcome variable, citizens' trust in policy, is measured in the survey using a composite index of six items (Cronbach’s alpha = `r alpha_air_trust` and `r alpha_waste_trust` for the air-pollution and waste policies respectively). This scale was adopted with modifications from validated trust scales used by previous public-administration studies [@karens_2016; @grimmelikhuijsen_2017; @grimmelikhuijsen_2018]. Participants were asked to indicate their agreement with the following statements, between 1 (weakly agree) and 7 (strongly agree): (1) I believe that the actions mentioned in the policy plan will assist in fulfilling the policy goal; (2) I believe that the actions mentioned in the policy plan were designed in a professional manner; (3) I believe that the policy plan is in the interest of citizens; (4) I believe that the policy plan reflects a genuine attempt to improve the well-being of citizens; (5) I believe that EPM made an honest attempt to design a good policy plan; (6) I believe that the Ministry of Environmental Protection aims to keep its commitments in that policy plan.

My key independent variables are the manipulations of symbols and information in the communication, as described above, and their interactions with the perceived personal relevance of the communication. Manipulation checks for the symbols confirm that the subjects in both areas are fairly familiar with the EPM symbols, and these symbols evoke positive feelings. Manipulation checks for the information confirm that the weak policy plans were perceived as less effective in achieving the policy goal compared with the strong plans. The perceived personal relevance of the communication is operationalized observationally (through the distinction between the Haifa-Bay and center samples) as well as experimentally (through the abovementioned priming questions). I coded living in Haifa-bay area according to respondents’ reporting on residence, work or studying in one of the Haifa-Bay cities. Validity checks confirm that the air-pollution policy is perceived as more personally relevant by Haifa Bay residents compared with center residents, whereas the recycling policy is perceived by both groups as equally relevant. Moreover, when directly asked about the most salient environmental issue in their area of residence, most of the Haifa-Bay subjects regarded to the air-pollution. As for the relevance priming questions, however, manipulation checks suggest that those in the relevance treatment group did not perceive either of the two policies as more personally relevant. This limitation should be kept in mind in the following analyses of the effect of this manipulation. The results of the all manipulation and validation checks  reported here are avilable from the supplementary appendix.

In addition to the above independent variables, the survey also included control items. Summary statistics and correlation matrix for all variables are available in the supplementary appendix.

@@

mod_air_1.1_r1 = lm(TRUST_air_INDEX~SYMBOL_t.r+INFORMATION_weak,data=SVIVA2_01_comb)

@@

**Table 1** presents the results of OLS regression analyses for the effects of the symbols and information manipulations on trust in the two policies, and their interaction with the areas. Models 1.1-1.3 regards the Air-pollution policy, and Models 1.4 - 1.6 regards the recycling policy. For each of these policy plans, I first regress trust on the manipulations (Models 1.1, 1.4), then I add the areas (Models 1.2, 1.5) and the interaction between them (Models 1.3, 1.6). The real symbols had a positive main effect on both policies compared with the no symbols condition. Compared with the fake symbols, the increase is significant at the 90% (one-tailed) regarding the air-pollution policy, yet it is insignificant compared with the recycling policy. The information manipulation has a significant main effect in both policy plans. As shown by the coefficients of the Area in Models 1.2 and 1.5, participants from Haifa-Bay reported lower levels of trust in the air-pollution policy, and also, to a lower degree, in the recycling policy. This may suggest that Haifa-Bay residents tend to have prior negative attitudes towards EPM's actions regarding the air-pollution in their area, but less so towards EPM or environmental policy in general.

I now turn to analyze the interactions between the manipulations and the areas. With regard to the air-pollution policy, my first hypotheses entails a positive interaction between residence in the center (i.e. low personal relevance) and the real symbols. My second hypothesis entails a positive interaction between center residents and weak information. Yet, counter to these expectations, in Model 1.3, the interaction between residence in the center and the real symbols is negative, and the interaction with the weak information is insignificant. These findings suggest that the real symbols were more effective on those reside in the polluted area and that the latter were not more likely to scrutinize the content of the policy plan. Similar patterns were also found with regard to the recycling policy. In Model 1.6, the interaction between the areas and the information is insignificant, and the effect of the real symbols is similarly greater among the Haifa-Bay subjects. **Figure 3** provides a descriptive demonstration of the effects of the the interaction between the areas and the symbols manipualation, to facilitate their interpretations.

@@

Thereafter, I will add the interaction with the relevance manipulation. In all these analyses I use linear OLS regressions.

@@

@@

>>>>

# Research design

@@

My outcome variable, citizens' trust in policy, is measured in the survey using a composite index of six items (Cronbach’s alpha = `r alpha_air_trust` and `r alpha_waste_trust` for the air-pollution and waste policies respectively). This scale was adopted with modifications from validated trust scales used by previous public-administration studies [@karens_2016; @grimmelikhuijsen_2017; @grimmelikhuijsen_2018]. Participants were asked to indicate their agreement with the following statements, between 1 (weakly agree) and 7 (strongly agree): (1) I believe that the actions mentioned in the policy plan will assist in fulfilling the policy goal; (2) I believe that the actions mentioned in the policy plan were designed in a professional manner; (3) I believe that the policy plan is in the interest of citizens; (4) I believe that the policy plan reflects a genuine attempt to improve the well-being of citizens; (5) I believe that EPM made an honest attempt to design a good policy plan; (6) I believe that the Ministry of Environmental Protection aims to keep its commitments in that policy plan. Participants were more likely to report low levels of trust in the air-pollution policy (*Mean* = `r SVIVA2_01_comb%>%summarise(mean(TRUST_air_INDEX,na.rm=T))%>%max() %>% round(2)`, *SD* = `r SVIVA2_01_comb%>%summarise(sd(TRUST_air_INDEX,na.rm=T))%>%max() %>% round(2)`) compared with the recycling policy (*Mean* = `r SVIVA2_01_comb%>%summarise(mean(TRUST_waste_INDEX,na.rm=T))%>%max() %>% round(2)`, *SD* = `r SVIVA2_01_comb%>%summarise(sd(TRUST_waste_INDEX,na.rm=T))%>%max() %>% round(2)`).

My key independent variables are the manipulations of symbols and information in the communication, as described above, and their interactions with the perceived personal relevance of the communication. Manipulation checks for the symbols confirm that the subjects in both areas are fairly familiar with the EPM symbols, and these symbols evoke positive feelings. Manipulation checks for the information confirm that the weak policy plans were perceived as less effective in achieving the policy goal compared with the strong plans. The perceived personal relevance of the communication is operationalized observationally (through the distinction between the two areas: Haifa-Bay and center) as well as experimentally (through the abovementioned relevance manipulation). I coded living in Haifa-bay area according to respondents’ reporting on residence, work or studying in one of the Haifa-Bay cities. Validity checks confirm that the air-pollution policy is perceived as more personally relevant by Haifa Bay residents compared with center residents, whereas the recycling policy is perceived by both groups as equally relevant. Moreover, when directly asked about the most salient environmental issue in their area of residence, most of the Haifa-Bay subjects regarded to the air-pollution. As for the relevance priming questions, however, manipulation checks suggest that those in the relevance treatment group did not perceive either of the two policies as more personally relevant. This limitation should be kept in mind when interpreting the results regarding this manipulation. The results of the all manipulation and validation checks are avilable from the supplementary appendix, along with summary statistics and correlation matrix for all variables.

@@

mod_air_1.1_r1 = lm(TRUST_air_INDEX~SYMBOL_t.r+INFORMATION_weak,data=SVIVA2_01_comb)
mod_waste_1.1_r1 = lm(TRUST_waste_INDEX~SYMBOL_t.r+INFORMATION_weak,data=SVIVA2_01_comb)

@@

**Table 1** presents the results of OLS regression analyses for the effects of the symbols and information manipulations on trust in the two policy plans, and their interaction with the two areas. Models 1.1-1.3 regard the Air-pollution policy, and Models 1.4 - 1.6 regard the recycling policy. I regress trust in policy on the manipulations (Models 1.1, 1.4), then I add the areas (Models 1.2, 1.5) and the interaction between them (Models 1.3, 1.6). The real symbols had positive main effects on both policies compared with the no symbols condition. Compared with the fake symbols, the increase is significant at the 90% (one-tailed) regarding the air-pollution policy, and only marginally significant compared with the recycling policy (*p* = `r (summary(mod_waste_1.1_r1)$coefficients["SYMBOL_t.r1",4]/2)%>%round(2)`, one-tailed). The information manipulation also has a significant main effect in both policy plans, as expected. Regarding the two areas, as shown by the coefficients of this factor in Models 1.2 and 1.5, participants from Haifa-Bay reported lower levels of trust in the air-pollution policy, and also, to a lower degree, in the recycling policy. The latter suggests that Haifa-Bay residents tend to have prior negative attitudes towards EPM's actions regarding the air-pollution in their area, and to some extent towards EPM or environmental policy in general.

Keeping in mind these main effects, I now turn to analyze the interactions between the manipulations and the areas. With regard to the air-pollution policy, my first hypotheses entails a positive interaction between residence in the center (i.e. low personal relevance) and the real symbols. Based on my second hypothesis, I expect to find a positive interaction between center residents and weak information. Counter to these expectations, in Model 1.3, the interaction between residence in the center and the real symbols is negative (*p*<.1), and the interaction with the weak information is insignificant. These findings suggest that the real symbols were more effective in increasing the trust of those reside in the polluted area and that the latter were not more likely to scrutinize the content of the communication. I also found similar patterns regarding the recycling policy. In Model 1.6, the interaction between the areas and the information is insignificant, and the interaction with the real symbols is negative. **Figure 3** provides a descriptive demonstration of the interaction between the areas and the symbols manipualation for the two policies.

@@

Next, I examine whether those subjects who were primed to think about the personal relevance of these policy plans were more likely to scrutinize them, and less likely to be affected by the familiar symbols. The interaction between the relevance manipulation and the symbols and information manipulations is presented in **Table 2**. Again, the first three models regard the Air-pollution policy, and the remaining models regard the recycling policy. Models 2.1, 2.4 replicate models 1.1, 1.4 while adding the relevance manipulation. Thereafter, I add the interactions between the relevance manipulation and the symbols and information manipulations (Models 2.2, 2.5). Finally, to test the moderation of the relevance manipulation within each area, I add three-way interactions with the areas (Models 2.3, 2.6). In all these models, the relevance treatment did not yield significant effects, and there are no significant interactions with the symbols and information manipulations. The triple interactions with the areas are also insignificant (in Models 2.3, 2.6), which indicates that in neither of these areas did the relevance treatment moderate the effects of the symbols and information manipulations.

@@

I further explored the robustness of these findings additional analyses, reported in the supplementary appendix. First, in addition to the above operationalizations of personal relevance, I also relied on subjects' subjective reporting. Second, I further explored the effect of personal relevance on subjects inclincation to scrutinize the policy plans, using additional measures for message elaboration available in the survey: the time participants spent on scanning the policy plans and their score in a memory test. I also tested wheather personal relevance decreased the delta (within-subjects) between trust in strong and weak policy plans presented to each subject. Third, I added a set of individual-level controls, and factors related to personal relevance of air-pollution, and to personal interest in environmental issues. Finally, I replicated my analyses on the unfilterred sample. None of these multiple robust analyses supports my theoretical hypotheses.

@@

Public administration studies have demonstrated the positive effect of symbolic elements in governemnt public communications on citizens trust in government organizations and their actions.
